---
title: "p8130_hw4_jsg2145"
author: "Jared Garfinkel"
date: "11/16/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = FALSE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

##Problem 3

```{r}
brain <- readxl::read_excel("./Brain.xlsx") %>% 
  janitor::clean_names()
```

```{r}
non_hum <- brain %>% 
  filter(species != "Homo sapiens")
```

```{r}
non_hum %>% 
  lm(glia_neuron_ratio ~ ln_brain_mass, data = .)
```

## Problem 3a

```{r}
non_hum %>% 
  ggplot(aes(x = ln_brain_mass, y = glia_neuron_ratio)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~x, se = FALSE)
```

```{r}
fit_1 <- non_hum %>% 
  lm(glia_neuron_ratio ~ ln_brain_mass, data = .)
```

```{r}
predict(fit_1, data.frame(ln_brain_mass = 7.22))
```
## Problem 3b

The predicted glia-neuron ratio for humans given the non human data is `r round(predict(fit_1, data.frame(ln_brain_mass = 7.22)), digits = 2)`.

## Problem 3c

The interval for the mean glia-neuron ratio at the given brain mass may make more 

sense given the data we have observed. An interval for a single new observation may

make more sense if we are interested in the predictive ability of our model.

## Problem 3d

$\hat{Y_h} = b_0 + b_1*X_h$

$$s^2(\hat{Y_h}) = MSE\left [\frac{1}{n} + \frac{(X_h-\bar{X})^2}{\sum{(X_i - \bar{X})^2}}\right]$$

```{r}
non_hum %>% 
  mutate(
    xbar = mean(ln_brain_mass),
    xdiff = ln_brain_mass - xbar,
    ybar = mean(glia_neuron_ratio),
    ydiff = glia_neuron_ratio - ybar) %>% 
  summarize(
    xdifftot = sum(xdiff^2),
    ydifftot = sum(ydiff^2),
    MSE = ydifftot/15
  )
```


$$s^2(\hat{Y_h}) = 0.0774*\left [ \frac{1}{17} + \frac{(7.22 - 4.29)^2}{22.21}\right]$$

= .0774 * `r 1/17 + ((7.22-4.29)^2/22.21)`

= `r .0774 * (1/17 + (7.22-4.29)^2/22.21)`

s = $\sqrt{s^2}$ = $\sqrt{.034} = `r sqrt(.034)`

95% CI: $$\hat{Y_h} \pm t_{1-\alpha/2,~n-2}*s(\hat{Y_H})$$

= 1.47 $\pm t_{.975,~15}*0.18$

= 1.47 $\pm$ `r qt(.975, 15)` * 0.18

= (`r 1.47 - qt(.975, 15)*0.18`,`r 1.47 + qt(.975, 15)*0.18`)

Since the human glia nueron ratio is within the 95% confidence interval of the predicted response

there is no evidence that the human brain has an excessive glia-neuron ratio compared with

other primates.

## Problem 3e

This is extrapolation since the prediction is outside the observed range. Extrapolation can 

provide accurate results when the prediction is close to the observed data. However, it may 

not be a good idea if the prediction is far from the observed data.